---
title: "Rserver"
author: "A.Bock"
date: "July 15, 2016"
output: pdf_document
---
This document is a summary of work and code being developed for the CDI project "Automated Mapping of SB OBjects"

This script will take the accessed SB data and GIS features, map them onto a leaflet map, and create several
summary graphs and downloadable files.  Note that the "shiny" app will be built upon this functionality.
Functions are a part of package specificed (*sbtools::item_get*  *item_get* is function in *sbtools* library)

## 1.  Plotting the Data

Import the necessary libraries
```{r}
library(leaflet)
library(RColorBrewer)
```

A couple of different ways to present color ramps to symbolize the segments with

```{r}
#pal<-colorNumeric(
#  palette = "Blues",
#  domain=sfMeans
#)

pal2 <- colorQuantile("YlGn", NULL, n = 10)
```

This next section determines what information is in the popup balloons when a user clicks on a feature/segment.
This needs to be changed to the stream GNIS name. (what else?)

```{r}
popup <- paste0("<strong>Name: </strong>", 
                GF_segs@data$seg_id)
```

Here is the map!
Note that we can offer an option to change the basemap (I believe)
Things to Do before carrying map over to shiny:
  1.  Add buffer to shapefile
  2.  Customize popup
  3.  Add legend to map
  4.  Set Bounds for map based on feature extent

```{r}
leaflet(GF_segs) %>% addTiles() %>% addPolylines(color=~pal2(sfMeans),weight=2,popup=~popup)
```